Relationships in the extremes and their influence on competition and coexistence

### 2021-09-14 paper/Paper.Rnw
continued editing to Intro

Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. %**JAS: should we introduce invader analysis or terminology here?
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.

%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
%Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
%While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
%
%***
%We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive.

%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.

%***DAN: stopped here 2021 07 20

>>>>

%Intro to storage effects and why they might be influenced by ATAs
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018} and we explain why ATAs may be relevant to them.
Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
whichever species is currently rare; this is the characteristic that separates so-called \emph{stabilizing} mechanisms of coexistence from
\emph{equalizing} mechanisms, which instead reduce average fitness differences between comeptitors \citep{Chesson_2000}.
Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive covariation between the quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance). 
%This is reasonable assumption for non-rare species, since 
%a better environment leads to greater numbers which leads to more intraspecific competition, but a rare species may be essentially unaffected
%by intraspecific competition. 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
in times when the environment is poor; environmentally insensitive adult stages and dormant stages such as seed banks are
example mechanisms \citep{Chesson_2012}. 
When present, these factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
Storage effects are quantified by measuring to what extent EC covariance contributes to differences between 
growth rates of species when they are at low ``invader'' densities, and growth rates of the abundant ``resident'' species \citep{Chesson_1994,Ellner_2016}. 
Thus ATAs may influence storage effects because storage effects involve EC covariation. ATAs between E and C may,
in principle, mediate the strength of storage effects, though to our knowledge this possibility has not previously been investigated.
%***
%Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
%In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
%The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
%In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
%In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. %**JAS: should we introduce invader analysis or terminology here?
%Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
%Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
%Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
%For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
%Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
%Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
